Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-08-04 paper-amia-summit-2018/manuscript.tex
update results section

Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers of electronic health records (EHR)\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection of EHR.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al. applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports in EHR into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al. trained Maximum Entropy models to classify the document into pre-defined sections, such as general patient information, provider information, medical condition, medical history, clinical history, findings, impression, etc.\cite{tepper2012statistical} Denny et al. proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision. \cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries in EHR. Liu et al. have shown that a linear chain CRF performed better than the hidden markov model and Maxent in sentence boundary detection, evaluated using the official
NIST evaluation tools.\cite{liu2005using} Recently, Treviso et al. proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word embeddings,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset.

In this study, we adopt the model proposed by Treviso et al. \cite{treviso2017sentence} and rename it as convolutional recurrent neural networks (CRNN) to apply it differently from the aforementioned studies. The main difference is that our study focuses on the segmentation of e-Coaching emails as distinct from determining the sentence boundary. Another important difference is that punctuation features were utilized in combination with word embeddings and part-of-speech features. Furthermore, we train our model with pre-trained word2vec word vectors,\cite{mikolov2013distributed} instead of using GloVe word vectors. Segmentation of e-Coaching emails is also different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.

@@

\section*{Results}
Our experimental results have two important directions. First, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}. Second, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features.

@@

@@

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together. Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.\\

>>>>

Relevant previous work in the biomedical domain primarily focused on segmentation of text into sections and headers of electronic health records (EHR)\cite{apostolova2009automatic,denny2009evaluation,tepper2012statistical,cho2002text} or sentence boundary detection of EHR.\cite{griffis2016quantitative,kreuzthaler2015detection,treviso2017sentence} Apostolova et al. applied SVM along with word-vector cosine similarity metric combined with several heuristics to segment clinical reports in EHR into sections, such as demographics, history, procedure, finding and impression.\cite{apostolova2009automatic} After identification of each line in the document, Tepper et al. trained Maximum Entropy models to classify the document into pre-defined sections, such as general patient information, provider information, medical condition, medical history, clinical history, findings, impression, etc.\cite{tepper2012statistical} Denny et al. proposed a SecTag algorithm, which combined natural language processing techniques, terminology-based rules and a Naive Bayes classifier to identify the sections and headers that achieved 99\% recall with 95.6\% precision. \cite{denny2009evaluation} On the other hand, SVM based on prosodic and part of speech features \cite{kreuzthaler2015detection} and recurrent convolutional neural networks using word embeddings \cite{griffis2016quantitative} were utilized for detecting sentence boundaries in EHR. Liu et al. have shown that a linear chain CRF performed better than the hidden markov model and Maxent in sentence boundary detection, evaluated using the official
NIST evaluation tools.\cite{liu2005using} Recently, Treviso et al. proposed a recurrent convolutional neural network for automatic identification of sentence boundary in a neuropsychological text in the Portuguese language.\cite{treviso2017sentence} Their proposed model combined prosodic and part of speech features with pre-trained GloVe word embeddings,\cite{pennington2014glove} which achieved 0.74 F1-score in sentence boundary detection on the Brazilian constitution dataset. Segmentation of e-Coaching emails is different from a traditional shallow discourse analysis of conversations \cite{galley2003discourse} in that the focus is on segmentation, rather than on determining the types of transitions between the utterances or assigning utterances to speakers.

@@

\section*{Results}
Our experimental results have four important directions. First, optimal sample size is identified by reporting MLP results with different sample size. Second, results are reported with respect to ``new segment'' class as well as weighted average over ``new segment'' and ``same segment'' classes in Table~\ref{tab:result_base}. Third, classification performance of different machine learning methods are outlined in Table~\ref{tab:result_weighted_avg} when word embeddings or lexical features are used in combination with punctuation and POS features. Fourth, the impact of individual fetures as well as their combination on the segmentation task are reported.

@@

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection when all features are utilized together. Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes when word embeddings or lexical features are used in combination with punctuation and POS features.

@@
